---
title: "Statistical Analysis Plan"
subtitle: |
  `r params$sap$study$study_code %||% ''`\
  `r if (is.na(params$sap$study$title %||% NA)) '' else params$sap$study$title`
date: "`r params$sap$study$date %||% Sys.Date()`"
author: "`r paste(unlist(params$sap$study$authors %||% ''), collapse = ', ')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    toc_depth: 3
    self_contained: true
  officedown::rdocx_document:
    toc: false
    number_sections: false
    reference_docx: reference.docx
    tables:
      style: Table
      caption:
        style: Table Caption
    page_margins:
      bottom: 1
      top: 1
      right: 1
      left: 1
      header: 0.5
      footer: 0.5
params:
  sap: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

`%||%` <- function(x, y) {
  if (is.null(x) || length(x) == 0) return(y)
  if (length(x) == 1 && is.na(x)) return(y)
  x
}

# The document renders the IncidencePrevalence calls the SAP describes, so it
# needs the generator that builds them (R/sap_code.R) and the cohort-kind helpers
# it stands on. Sourced rather than copied: the SAP and the code it generates must
# never be able to drift apart, which is the whole reason this section exists.
#
# knitr runs with the working directory set to this file's own, so R/ is one
# level up; the second candidate lets the document render from the repo root too.
# local = knit_global() puts the definitions where the chunks can see them,
# instead of inside this helper's frame.
source_app <- function(file) {
  cand <- c(file.path("..", "R", file), file.path("R", file))
  hit  <- cand[file.exists(cand)]
  if (!length(hit)) {
    stop("sap_preview.Rmd cannot find R/", file, " (working directory: ", getwd(), ")")
  }
  source(hit[[1]], local = knitr::knit_global())
}
for (f in c("utils.R", "cohort_kinds.R", "analysis_registry.R", "sap_code.R",
            "cohort_operations.R")) source_app(f)

is_word <- isTRUE(knitr::opts_knit$get("rmarkdown.pandoc.to") == "docx")

# A fenced R block. Word gets the same fence: pandoc's docx writer renders it as
# a monospaced source-code paragraph, which is what the reference SAPs look like.
code_block <- function(code) {
  if (is.null(code) || !nzchar(code)) return(invisible(NULL))
  cat("\n``` r\n", code, "\n```\n\n", sep = "")
}

sap   <- params$sap
study <- sap$study %||% list()

library(flextable)
set_flextable_defaults(font.family = "Calibri", font.size = 10)

# Render a two-column label/value flextable, skipping rows with no value
field_table <- function(rows) {
  rows <- Filter(function(r) {
    v <- r[2]
    !is.null(v) && !is.na(v) && nzchar(trimws(v)) && v != "—"
  }, rows)
  if (!length(rows)) return(invisible(NULL))
  df <- as.data.frame(do.call(rbind, rows), stringsAsFactors = FALSE)
  colnames(df) <- c("Field", "Value")
  ft <- flextable(df) |>
    bold(j = 1) |>
    border_remove() |>
    hline(border = officer::fp_border(color = "#DDDDDD", width = 0.5)) |>
    set_table_properties(layout = "autofit", width = 1) |>
    padding(padding = 4) |>
    fontsize(size = 10)
  if (nrow(df) >= 2) ft <- bg(ft, i = seq(2, nrow(df), 2), bg = "#F5F5F5")
  ft
}

# Render a data frame as a flextable, one row per item (sources, changes)
items_table <- function(df) {
  ft <- flextable(df) |>
    bold(part = "header") |>
    border_remove() |>
    hline(border = officer::fp_border(color = "#DDDDDD", width = 0.5)) |>
    set_table_properties(layout = "autofit", width = 1) |>
    padding(padding = 4) |>
    fontsize(size = 10)
  if (nrow(df) >= 2) ft <- bg(ft, i = seq(2, nrow(df), 2), bg = "#F5F5F5")
  ft
}

# A field as display text: NULL and NA become the empty string
chr <- function(x) {
  x <- as.character(x %||% "")
  if (is.na(x)) "" else x
}

# Every CDM source key the SAP defines. When an item runs on all of them --
# the usual case -- its Data sources cell says so in two words instead of
# repeating the same five keys on every row.
all_source_keys <- vapply(sap$cdm_sources %||% list(), function(s) {
  key <- chr(s$source_key)
  if (nzchar(key)) key else chr(s$name)
}, character(1))
sources_cell <- function(x) {
  ds <- as.character(unlist(x$data_sources %||% list()))
  if (length(ds) && length(all_source_keys) && setequal(ds, all_source_keys)) {
    return("All data sources")
  }
  paste(ds, collapse = ", ")
}

```

```{r justify-html, results='asis'}
# The Word output justifies body text via reference.docx; mirror it in HTML.
# li included: the numbered objectives are a list, not paragraphs.
if (!is_word) cat("<style>p, li { text-align: justify; }</style>\n")
```

# Study information

```{r study-table}
rows <- list(
  c("Title",       study$title %||% ""),
  c("Study code",  study$study_code %||% ""),
  c("Authors",     paste(unlist(study$authors %||% list()), collapse = ", ")),
  c("SAP version", study$version %||% ""),
  c("Date",        study$date %||% ""),
  # Stated even though it is usually the conventional 5: a reader must be able to
  # see the export threshold without inferring it, and a plan that states none is
  # a different claim from one that states 5.
  c("Minimum cell count",
    if (is.null(study$min_cell_count %||% NULL)) "Not stated"
    else sprintf("%s (results below this are suppressed before export)",
                 format(study$min_cell_count)))
)
field_table(rows)
```

```{r amendments, results='asis'}
amds <- study$amendments %||% list()
if (length(amds)) {
  cat("\n**Amendment history**\n\n")
  df <- data.frame(
    `Version`               = vapply(amds, function(a) chr(a$version), character(1)),
    `Date`                  = vapply(amds, function(a) chr(a$date), character(1)),
    `Description of change` = vapply(amds, function(a) chr(a$description), character(1)),
    check.names = FALSE, stringsAsFactors = FALSE
  )
  flextable::flextable_to_rmd(items_table(df))
  cat("\n\n")
}
```

```{r background, results='asis'}
bg_text <- study$background %||% ""
if (!is.na(bg_text) && nzchar(trimws(bg_text))) {
  cat("\n**Rationale and background**\n\n", bg_text, "\n\n", sep = "")
}
```

```{r objectives, results='asis'}
aim  <- chr(study$aim)
objs <- vapply(study$objectives %||% list(), objective_text, character(1))
if (nzchar(trimws(aim)) || length(objs)) {
  cat("\n**Research question and objectives**\n\n")
  if (nzchar(trimws(aim))) cat(aim, "\n\n")
  if (length(objs)) {
    cat("The specific objectives of this study are:\n\n")
    cat(sprintf("%d. %s\n", seq_along(objs), objs), sep = "")
  }
}
```

`r if (is_word) "\\newpage"`

---

# CDM sources

```{r cdm-sources, results='asis'}
sources <- sap$cdm_sources %||% list()
if (!length(sources)) {
  cat("_No CDM sources defined._\n")
} else {
  df <- data.frame(
    `Source name`      = vapply(sources, function(s) chr(s$name), character(1)),
    `Short key`        = vapply(sources, function(s) chr(s$source_key), character(1)),
    `Country / region` = vapply(sources, function(s) chr(s$country), character(1)),
    check.names = FALSE, stringsAsFactors = FALSE
  )
  flextable::flextable_to_rmd(items_table(df))
  cat("\n\n")
}
```

`r if (is_word) "\\newpage"`

---

# CDM changes

```{r cdm-changes, results='asis'}
changes <- sap$cdm_changes %||% list()
if (!length(changes)) {
  cat("_No CDM changes defined._\n")
} else {
  df <- data.frame(
    `Type of change` = vapply(changes, function(ch) chr(ch$change_type), character(1)),
    `Data sources`   = vapply(changes, function(ch)
      paste(unlist(ch$data_sources %||% list()), collapse = ", "), character(1)),
    `Description`    = vapply(changes, function(ch) chr(ch$description), character(1)),
    check.names = FALSE, stringsAsFactors = FALSE
  )
  flextable::flextable_to_rmd(items_table(df))
  cat("\n\n")
}
```

`r if (is_word) "\\newpage"`

---

# Codelists

```{r codelists, results='asis'}
codelists <- sap$codelists %||% list()
if (!length(codelists)) {
  cat("_No codelists defined._\n")
} else {
  # The codes themselves are NOT in this table. Listing them one per line inside
  # a single cell made one row 41 lines tall for a 75-code list, and still cut it
  # off at 40 -- so the section was both bulky and incomplete. They now live in
  # full, untruncated, in the codelist appendix; this cell says how many to expect.
  #
  # Counted from the EXPRESSION, which is what the author chose. An expression
  # that pulls in descendants resolves to more concepts than it names, so the
  # count is stated as a floor rather than as a total the SAP cannot know without
  # a vocabulary to resolve against.
  codes_cell <- function(cl) {
    expr <- cl$concept_set_expression %||% list()
    if (!length(expr)) return("— not yet uploaded")
    if (concept_set_expands(expr)) {
      return(sprintf("%d named, plus descendants (listed in the appendix)", length(expr)))
    }
    sprintf("%d (listed in full in the appendix)", length(expr))
  }
  describe <- function(cl) {
    d   <- chr(cl$description)
    src <- chr(cl$source_file)
    if (nzchar(src)) d <- if (nzchar(d)) sprintf("%s (from %s)", d, src) else sprintf("from %s", src)
    d
  }
  # Grouped by category into one sub-table each, the way DARWIN SAPs lay out
  # their codelist appendix (Index events, Comorbidities, ...). An uncategorised
  # codelist falls into "Other"; the canonical order leads, then any free-text
  # categories, then Other last.
  cats <- vapply(codelists, function(cl) {
    c <- chr(cl$category)
    if (nzchar(c)) c else "Other"
  }, character(1))
  # Canonical categories, in the order the DARWIN appendix uses, each with its
  # (irregular) plural heading. Free-text categories keep their own name.
  canon    <- c("Index event", "Medicine", "Procedure", "Condition / observation",
                "Comorbidity", "Outcome")
  headings <- c(`Index event` = "Index events", Medicine = "Medicines",
                Procedure = "Procedures", `Condition / observation` = "Conditions / observations",
                Comorbidity = "Comorbidities", Outcome = "Outcomes")
  ordered <- c(intersect(canon, cats), setdiff(unique(cats), c(canon, "Other")), "Other")
  for (cat in ordered[ordered %in% cats]) {
    idx     <- which(cats == cat)
    heading <- if (cat %in% names(headings)) headings[[cat]] else cat
    cat("\n\n## ", heading, "\n\n", sep = "")
    df <- data.frame(
      `Codelist`    = vapply(codelists[idx], function(cl) chr(cl$name), character(1)),
      `Description` = vapply(codelists[idx], describe, character(1)),
      `Concepts`    = vapply(codelists[idx], codes_cell, character(1)),
      check.names = FALSE, stringsAsFactors = FALSE
    )
    flextable::flextable_to_rmd(items_table(df))
    cat("\n\n")
  }
}
```

`r if (is_word) "\\newpage"`

---

# Description of cohorts

```{r cohorts, results='asis'}
cohorts <- sap$cohorts %||% list()
if (!length(cohorts)) {
  cat("_No cohorts defined._\n")
} else {
  # A generated denominator has no entry logic -- its "logic" IS the generator
  # arguments, kept as the compact facts line.
  denominator_facts <- function(co) {
    parts <- character(0)
    add <- function(label, value) {
      if (nzchar(value)) parts <<- c(parts, sprintf("%s: %s", label, value))
    }
    if (!is.null(co$targetCohortTable))
      add("Target cohort", chr(co$targetCohortTable))
    if (!is.null(co$daysPriorObservation))
      add("Days prior observation", paste(unlist(co$daysPriorObservation), collapse = ", "))
    if (!is.null(co$ageGroup))
      add("Age groups", paste(
        vapply(co$ageGroup, function(a) paste(unlist(a), collapse = "–"), character(1)),
        collapse = "; "))
    sex <- unlist(co$sex %||% list())
    if (length(sex)) add("Sex", paste(sex, collapse = ", "))
    if (!is.null(co$cohortDateRange)) {
      # An absent bound is NOT missing information: NULL on either end means the
      # cohort runs to the limit of what the database holds, which is what the
      # generated call's NA does. It used to print as an em-dash -- the same
      # glyph field_table() uses for "no value worth showing" -- so a deliberate
      # open end and an unanswered question read identically.
      #
      # Indexed with [[ ]] and NOT unlist()ed: unlist drops a NULL rather than
      # preserving its position, so an open START slid into position 1 and a
      # range of [null, "2020-01-01"] printed as "2020-01-01 to <open>" -- the
      # opposite period to the one the SAP states. Position is the whole meaning
      # of this pair, so it has to survive.
      dr    <- co$cohortDateRange
      bound <- function(i, whole) {
        v <- if (length(dr) >= i) dr[[i]] else NULL
        v <- as.character(v %||% NA)[1]
        if (is.na(v) || !nzchar(trimws(v))) whole else v
      }
      add("Cohort date range", sprintf("%s to %s",
                                       bound(1, "start of available data"),
                                       bound(2, "end of available data")))
    }
    tar_list <- co$timeAtRisk %||% list()
    if (length(tar_list))
      add("Time at risk", paste(vapply(tar_list, function(w) {
        w <- unlist(w)
        hi <- if (is.null(w[2]) || is.na(w[2]) || is.infinite(as.numeric(w[2]))) "∞" else w[2]
        sprintf("[%s, %s]", w[1], hi)
      }, character(1)), collapse = "; "))

    # The RULE that combines the axes above, and so how many cohorts the card
    # generates. Without it the age / sex / prior-observation lists are ambiguous:
    # 4 age groups and 2 sexes is 8 cohorts with interactions and 5 without. The
    # default matches denominator_cohort_set(), so a legacy file reads the same
    # way here as it does everywhere else.
    add("Requirements combined", if (isTRUE(co$requirementInteractions %||% TRUE)) {
      "every combination (requirementInteractions = TRUE)"
    } else {
      "each factor against the first value of the others, so order matters (requirementInteractions = FALSE)"
    })
    if (identical(canonical_cohort_kind(co$kind), "target_denominator")) {
      add("Requirements at entry", if (isTRUE(co$requirementsAtEntry)) {
        "must be met on the target cohort start date (requirementsAtEntry = TRUE)"
      } else {
        "may be met part-way through time in the target cohort (requirementsAtEntry = FALSE)"
      })
    }
    add("Cohorts generated", sprintf("%d", length(denominator_cohort_set(co))))

    paste(parts, collapse = "\n")
  }

  # A plain cohort's logic, IN ORDER, numbered: the entry event is step 1 --
  # it is the index event -- then the inclusion / exclusion criteria; exit on
  # a final line. flextable turns \n into real line breaks in HTML and Word.
  # Entry and inclusion are ONE ordered sequence -- the numbering runs straight
  # through -- but they are not the same kind of step, and until 0.4.19 this cell
  # ran them together unlabelled so the difference was invisible in the document.
  #
  # The entry event is the INDEX event: it decides the cohort start date. An
  # inclusion criterion filters people who already have one. A reader who cannot
  # tell which line set the index date cannot check the cohort definition at all,
  # and the two are separate fields on the card precisely because they are
  # separate operations.
  # Typed operations win when a cohort has them, and the sentences are GENERATED
  # from them (cohort_operations_prose) rather than written beside them. That is
  # the inversion the operations exist for: one source, so the logic a reviewer
  # reads and the code the study runs cannot describe different cohorts. A cohort
  # authored as free text is unaffected and renders exactly as before.
  cohort_logic <- function(co) {
    ops <- cohort_operations_prose(co)
    if (length(ops)) {
      return(paste(c("Generated from typed steps, in order:",
                     sprintf("%d. %s", seq_along(ops), ops)), collapse = "\n"))
    }
    entry <- unlist(co$entry_events %||% list())
    incl  <- unlist(co$inclusion_criteria %||% list())
    exits <- unlist(co$exit_criteria %||% list())
    txt   <- character(0)
    if (length(entry)) {
      txt <- c(txt, "Entry:", sprintf("%d. %s", seq_along(entry), entry))
    }
    if (length(incl)) {
      txt <- c(txt, "Inclusion / exclusion:",
               sprintf("%d. %s", length(entry) + seq_along(incl), incl))
    }
    # Exit criteria are bulleted, not numbered: they are alternatives that end an
    # episode, not further steps in the sequence above. No "earliest of" gloss --
    # these are the author's free-text lines, so let their wording carry it.
    if (length(exits)) txt <- c(txt, "Exit:", paste0("• ", exits))
    if (!length(txt)) return(denominator_facts(co))
    paste(txt, collapse = "\n")
  }

  # The same cell, as bold-labelled chunks instead of one flat string.
  #
  # A flextable cell is plain text unless it is COMPOSED from chunks, so styling
  # only the three labels means building the cell piece by piece and writing it
  # back with compose(). Kept beside cohort_logic() rather than replacing it: the
  # plain version still decides what a denominator cohort shows (its facts), and
  # having both means the styled and unstyled forms cannot say different things.
  logic_paragraph <- function(co) {
    ops <- cohort_operations_prose(co)
    if (length(ops)) {
      return(as_paragraph(
        as_chunk("Generated from typed steps, in order:",
                 props = fp_text_default(bold = TRUE)),
        as_chunk(paste0("\n", paste(sprintf("%d. %s", seq_along(ops), ops),
                                    collapse = "\n")))))
    }
    entry <- unlist(co$entry_events %||% list())
    incl  <- unlist(co$inclusion_criteria %||% list())
    exits <- unlist(co$exit_criteria %||% list())
    if (!length(entry) && !length(incl) && !length(exits)) {
      return(as_paragraph(as_chunk(denominator_facts(co))))
    }
    chunks <- list()
    section <- function(label, lines) {
      if (!length(lines)) return()
      lead <- if (length(chunks)) "\n" else ""
      chunks[[length(chunks) + 1]] <<-
        as_chunk(paste0(lead, label), props = fp_text_default(bold = TRUE))
      chunks[[length(chunks) + 1]] <<- as_chunk(paste0("\n", paste(lines, collapse = "\n")))
    }
    section("Entry:", sprintf("%d. %s", seq_along(entry), entry))
    section("Inclusion / exclusion:",
            sprintf("%d. %s", length(entry) + seq_along(incl), incl))
    section("Exit:", paste0("• ", exits))
    do.call(as_paragraph, chunks)
  }

  # The [bracketed] codelist references cited anywhere in the cohort's text --
  # the convention exists precisely so this column can be derived, not retyped.
  codelists_used <- function(co) {
    txt  <- c(unlist(co$entry_events %||% list()),
              unlist(co$inclusion_criteria %||% list()),
              unlist(co$exit_criteria %||% list()))
    if (!length(txt)) return("")
    refs <- unlist(regmatches(txt, gregexpr("\\[[^][]+\\]", txt)))
    paste(unique(gsub("^\\[|\\]$", "", refs)), collapse = "\n")
  }

  df <- data.frame(
    `Cohort`           = vapply(cohorts, function(co) chr(co$name), character(1)),
    `Kind`             = vapply(cohorts, function(co) chr(co$kind), character(1)),
    `Data sources`     = vapply(cohorts, sources_cell, character(1)),
    `Codelists used`   = vapply(cohorts, codelists_used, character(1)),
    `Logic (in order)` = vapply(cohorts, cohort_logic, character(1)),
    check.names = FALSE, stringsAsFactors = FALSE
  )
  # compose() takes one row at a time, because each cell has a different number of
  # chunks -- as_paragraph() on its own is vectorised across rows and cannot vary
  # its structure per row.
  ft <- items_table(df)
  for (i in seq_along(cohorts)) {
    ft <- compose(ft, i = i, j = "Logic (in order)", value = logic_paragraph(cohorts[[i]]))
  }
  flextable::flextable_to_rmd(ft)
  cat("\n\n")
}
```

`r if (is_word) "\\newpage"`

---

# Analyses

```{r analyses, results='asis'}
analyses <- sap$proposed_analyses %||% list()
if (!length(analyses)) {
  cat("_No analyses defined._\n")
} else {
  analysis_details <- function(a) {
    p <- a$parameters %||% list()
    parts <- character(0)
    add <- function(label, value) {
      if (nzchar(value)) parts <<- c(parts, sprintf("%s: %s", label, value))
    }

    # Cohort NAMES only. The *CohortId keys used to be a fallback here, which
    # would have labelled a numeric id array as if it were the cohort's name --
    # the ids say which cohorts OF A SET to use, and get their own entry below.
    add("Denominator", chr(p$denominatorTable %||% p$denominator_cohort))
    add("Outcome",     chr(p$outcomeTable %||% p$outcome_cohort))
    add("Censoring",   chr(p$censorTable))

    # Absent means "every cohort in the set" -- the estimators' own default -- so
    # only a RESTRICTED selection is worth stating.
    ids <- function(label, x) {
      v <- suppressWarnings(as.numeric(unlist(x %||% list())))
      v <- v[!is.na(v)]
      if (length(v)) add(label, paste(v, collapse = ", "))
    }
    ids("Denominator cohort IDs", p$denominatorCohortId)
    ids("Outcome cohort IDs",     p$outcomeCohortId)
    ids("Censor cohort IDs",      p$censorCohortId)

    if (!is.null(p[["interval"]]))
      add("Interval", paste(unlist(p[["interval"]]), collapse = ", "))

    if (!is.null(p$timePoint))
      add("Time point", paste(unlist(p$timePoint), collapse = ", "))

    # Period prevalence only. Both change WHO is counted in an interval, so a SAP
    # that omitted them read identically to one run with the opposite settings.
    if (!is.null(p$fullContribution))
      add("Full interval contribution required", if (isTRUE(p$fullContribution)) "yes" else "no")
    if (!is.null(p$completeDatabaseIntervals))
      add("Complete database intervals only", if (isTRUE(p$completeDatabaseIntervals)) "yes" else "no")
    if (!is.null(p$level))
      add("Estimation level", chr(p$level))

    # Incidence only.
    w <- washout_days(p$outcomeWashout)
    if (!is.null(w))
      add("Outcome washout", if (is.infinite(w)) "unbounded" else sprintf("%s days", format(w)))
    if (!is.null(p$repeatedEvents))
      add("Repeated events", if (isTRUE(p$repeatedEvents)) "counted" else "first event only")

    # `strata` is a list of variable GROUPS, and the grouping is the meaning:
    # list("sex", "age_group") is two separate stratifications, while
    # list(c("sex", "age_group")) is a single stratification by their cross. The
    # old unlist() here flattened both to the same string, so two materially
    # different analyses previewed identically. Crossed variables join with ×;
    # separate stratifications are listed one per line.
    groups <- p$strata %||% list()
    if (length(groups)) {
      add("Strata", paste(vapply(groups, function(g)
        paste(as.character(unlist(g)), collapse = " × "), character(1)), collapse = "\n"))
      # Inert without strata, so it is stated only where it does something.
      add("Overall (unstratified) result", if (isTRUE(p$includeOverallStrata)) "also reported" else "not reported")
    }

    paste(parts, collapse = "\n")
  }

  # This table is the whole of the Analyses section: every argument the author
  # decided, one row per analysis. The generated calls are NOT repeated here --
  # all generated code lives in the analysis-code appendix, which carries each
  # call under its own `# <analysis name>` comment, including the fallback note
  # for an analysis type no IncidencePrevalence function maps onto.
  df <- data.frame(
    `Name`          = vapply(analyses, function(a) chr(a$name), character(1)),
    `Analysis type` = vapply(analyses, function(a) chr(a$analysis_type), character(1)),
    `Data sources`  = vapply(analyses, sources_cell, character(1)),
    `Details`       = vapply(analyses, analysis_details, character(1)),
    check.names = FALSE, stringsAsFactors = FALSE
  )
  flextable::flextable_to_rmd(items_table(df))
  cat("\n\n")

  # Which analyses answer which objective. The plan states objectives at the top
  # and analyses in the middle, and until 0.4.22 nothing joined the two -- a
  # reviewer checking that the study answers what it set out to answer had to do
  # it by eye. Derived from the analyses' own objective references, so it cannot
  # disagree with them.
  objs <- study$objectives %||% list()
  if (length(objs)) {
    cat("## Objectives answered\n\n")
    answered <- function(i) {
      id <- objective_id(objs[[i]])
      nm <- vapply(analyses, function(a) {
        if (id %in% as.character(unlist(a$objectives %||% list()))) chr(a$name) else ""
      }, character(1))
      nm <- nm[nzchar(nm)]
      if (!length(nm)) "— no analysis yet" else paste(nm, collapse = "\n")
    }
    df <- data.frame(
      `#`         = as.character(seq_along(objs)),
      `Objective` = vapply(objs, objective_text, character(1)),
      `Analyses`  = vapply(seq_along(objs), answered, character(1)),
      check.names = FALSE, stringsAsFactors = FALSE
    )
    flextable::flextable_to_rmd(items_table(df))
    cat("\n\n")
  }
}
```

`r if (is_word) "\\newpage"`

---

# Appendix: codelist contents

```{r codelist-contents, results='asis'}
# Every code the SAP carries, in full and untruncated -- the summary table above
# gives only the counts. Here rather than beside the summary because this is bulk
# a reader consults, not reads: the same reason DARWIN SAPs put their codelists in
# an appendix. Nothing is repeated between the two.
#
# EVERY codelist gets a heading, including one with no codes yet. Filtering the
# empty ones out left a plan naming six codelists with one appendix entry, and
# the summary table promises each is "listed in full in the appendix" -- so a
# reader had to notice an absence to learn the codes were missing. An empty
# codelist says so under its own heading instead, the same way an analysis with
# no package call still gets a block in the analysis-code appendix.
all_codelists <- sap$codelists %||% list()

# One codelist as the densest table its codes allow. Concept ids alone tile
# several to a row; named concepts need a line each, because an OMOP concept name
# runs to sixty-odd characters and even two per row would overrun the page.
codes_table <- function(codes, per_row = 5) {
  code <- vapply(codes, function(cd) chr(cd$code), character(1))
  name <- vapply(codes, function(cd) chr(cd$name), character(1))
  if (any(nzchar(name))) {
    return(data.frame(`Concept ID` = code, `Concept name` = name,
                      check.names = FALSE, stringsAsFactors = FALSE))
  }
  rows <- ceiling(length(code) / per_row)
  grid <- matrix(c(code, rep("", rows * per_row - length(code))),
                 nrow = rows, byrow = TRUE)
  df <- as.data.frame(grid, stringsAsFactors = FALSE)
  names(df) <- rep("Concept IDs", per_row)
  df
}

# Smaller and tighter than items_table(): a code list is reference material, read
# by lookup rather than straight through, so it earns a denser setting than the
# tables in the body.
codes_ft <- function(df) {
  ft <- flextable(df) |>
    bold(part = "header") |>
    border_remove() |>
    hline(border = officer::fp_border(color = "#DDDDDD", width = 0.5)) |>
    set_table_properties(layout = "autofit", width = 1) |>
    padding(padding = 1.5) |>
    # part = "all": fontsize() defaults to the body, which would leave a 10pt
    # header sitting above 8pt rows.
    fontsize(size = 8, part = "all")
  if (nrow(df) >= 2) ft <- bg(ft, i = seq(2, nrow(df), 2), bg = "#F5F5F5")
  ft
}

# An EXPANDING expression is shown as the expression, flags and all, because
# then the flags are the meaning: "include descendants" is the difference
# between one concept and a whole subtree, and a reader cannot check the
# codelist without seeing which concepts carry it. A flat expression has nothing
# to show but its ids, so it keeps the denser tiled table.
expression_table <- function(expr, codes) {
  named <- vapply(codes, function(cd) chr(cd$name), character(1))
  names(named) <- vapply(codes, function(cd) chr(cd$code), character(1))
  ids <- vapply(expr, function(e) chr(e$concept_id), character(1))
  nm  <- unname(named[ids])
  nm[is.na(nm)] <- ""
  yn  <- function(f) vapply(expr, function(e) if (isTRUE(e[[f]])) "yes" else "", character(1))
  data.frame(
    `Concept ID`   = ids,
    `Concept name` = nm,
    `Descendants`  = yn("descendants"),
    `Mapped`       = yn("mapped"),
    `Excluded`     = yn("excluded"),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}

if (!length(all_codelists)) {
  cat("_No codelists defined._\n")
} else {
  for (cl in all_codelists) {
    expr  <- cl$concept_set_expression %||% list()
    codes <- cl$codes %||% list()
    cat("\n\n## ", chr(cl$name), "\n\n", sep = "")
    if (!length(expr) && !length(codes)) {
      # Stated, not implied by an empty section: this is the plan admitting the
      # codes are still to come, which is a reviewable fact about the SAP.
      cat("_No codes uploaded yet._\n\n")
      next
    }
    src   <- chr(cl$source_file)
    vocab <- chr(cl$vocabulary_version)
    from  <- if (nzchar(src)) sprintf(", from `%s`", src) else ""
    if (concept_set_expands(expr)) {
      n_desc <- sum(vapply(expr, function(e) isTRUE(e$descendants), logical(1)))
      n_excl <- sum(vapply(expr, function(e) isTRUE(e$excluded), logical(1)))
      cat(sprintf(paste("%d concept%s named%s%s%s. This is a concept set",
                        "expression, not a fixed list: the codelist it resolves to",
                        "is produced against each data source's own vocabulary when",
                        "the study runs.\n\n"),
                  length(expr), if (length(expr) == 1) "" else "s",
                  if (n_desc) sprintf(", %d including descendants", n_desc) else "",
                  if (n_excl) sprintf(", %d excluded", n_excl) else "", from))
      flextable::flextable_to_rmd(codes_ft(expression_table(expr, codes)))
      cat("\n\n")
      next
    }
    # A flat expression resolves to exactly the concepts it names, so the
    # snapshot IS the codelist -- and the vocabulary it came from is worth
    # stating, since another version may resolve the same ids differently.
    cat(sprintf("%d concept%s%s.%s\n\n", length(codes),
                if (length(codes) == 1) "" else "s", from,
                if (nzchar(vocab)) sprintf(" Vocabulary version: %s.", vocab) else ""))
    if (!length(codes)) {
      cat("_Every concept named is excluded, so this codelist resolves to nothing._\n\n")
      next
    }
    flextable::flextable_to_rmd(codes_ft(codes_table(codes)))
    cat("\n\n")
  }
}
```

`r if (is_word) "\\newpage"`

---

# Appendix: analysis code

```{r script, results='asis'}
# The whole plan as code, in the order it has to run: every denominator cohort
# set generated before the estimates that point at it. This is the executable
# form of the document, and the ONLY place generated code appears -- the sections
# above describe the same analyses in words and arguments.
#
# One heading per block rather than one flat listing. The body of the document
# names every cohort and analysis, so an appendix that ran them together left the
# reader checking a specific analysis's call with nowhere to jump to and a single
# long block to scan. The headings put each one in the table of contents; running
# the plan still means running the blocks in the order they appear.
#
# `cdm` is the live database connection the study opens; every other argument
# comes from this SAP. An argument the plan does not decide is absent, and the
# package's documented default applies.
secs <- sap_script_sections(sap)
if (!length(secs)) {
  cat("_No cohorts or analyses map onto IncidencePrevalence functions yet._\n")
} else {
  cat("Plain cohorts (target, outcome, comparator, censoring) are instantiated",
      "outside IncidencePrevalence and are referred to here by table name only.",
      "The blocks below run in the order they appear.\n\n")
  group <- NULL
  for (s in secs) {
    if (!identical(s$group, group)) {
      group <- s$group
      cat("\n\n## ", group, "\n\n", sep = "")
    }
    if (!is.na(s$title)) cat("### ", s$title, "\n\n", sep = "")
    if (is.null(s$code)) cat("_", s$note, "_\n\n", sep = "") else code_block(s$code)
  }
}
```

---

`r sprintf("_Generated by shinySAP · schema %s · %s_", sap$sap_schema_version %||% "unknown", sap$generated_at %||% as.character(Sys.time()))`
